\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
\usepackage{hyperref}

%SetFonts
%SetFonts

\title{Quantum Information Class 2024}
\author{Geza Toth}
\date{\today}							% Activate to display a given date or no date

\begin{document}
\maketitle
%\section{}
%\subsection{}

\begin{center}
See  \href{https://www.gtoth.eu/courses.html}{https://www.gtoth.eu/courses.html}
\end{center}

\begin{itemize}

\item Introduction

\item General characteristics of multi-partite quantum systems
\begin{itemize}
\item Statistics of classical bits, probabilistic logic by von Neumann
\item Quantum bits - a comparison with the statistics of classical bits
\item Geometry of quantum quantum states
\begin{itemize}
\item Bloch vector for $d=2$ (a single qubits)
\item Bloch vector for $d>2$
\end{itemize}
\item Convexity of the set of quantum states 
\begin{itemize}
\item Quantum states with full rank are inside the set, with derivation
\item Quantum states with non-full rank are on the boundary, with derivation
\end{itemize}
\item Reduced states, partial trace
\item Schmidt decomposition for bipartite pure states
\item Purification
\item Entropy
\begin{itemize}
\item Shannon entropy
\item Von Neumann entropy, main properties.
\begin{itemize}
\item Prove concavity with Klein’s inequality
\item prove that it is additive for independent systems
\item Prove subadditivity.
\item Araki-Lieb inequality (mention)
\item Strongly subadditive (mention)
 \end{itemize} 
\item Quantum conditional entropy
\item Quantum mutual information
\item Quantum relative entropy, properties, including monotonicity under CP maps.
\item Linear entropy, relation to von Neumann entropy via the Mercator series
\end{itemize} 
\item Quantum Fidelity and its properties
\item Distances (Bures distance, Hilbert-Schmidt distance, Trace distance)
\end{itemize}

\item Interesting quantum states
\begin{itemize}
\item Single-particle states
\item Bipartite singlet state, invariance under $U \otimes U$
\item Werner state, invariance under $U \otimes U$, full set
\item Isotropic state. invariance under $U \otimes U^*$
\item Schrödinger cat state
\item GHZ state and its stabilizer group (xxx,zz1,z1z,1zz,etc.)
\item W state
\item Symmetric Dicke state and its defining properties
\end{itemize}

\item Bell inequalities
\begin{itemize}
\item EPR paradox
\item LHV models
\item CHSH inequality
\item Mermin's inequality
\item Loopholes (Detection efficiency loophole, locality loophole)
\end{itemize}

\item Paper review: Gregor Weihs, Thomas Jennewein, Christoph Simon, Harald Weinfurter, and Anton Zeilinger, Violation of Bell’s Inequality under Strict Einstein Locality Conditions, Phys. Rev. Lett. 81, 5039  (1998)

\item Paper review: J. Yin, Y. Cao, Y.-H. Li et al., Satellite-based entanglement distribution over 1200 kilometers, Science 356, 1140-1144 (2017)

\item Entanglement theory (entangled/not entangled)
\begin{itemize}
\item Bipartite case
\begin{itemize}
\item Pure states
\item Mixed states (Werner  1989)
\end{itemize}
\item Entanglement criteria
\begin{itemize}
\item Partial transposition, with derivation 
\item Entanglement witnesses, several examples, e.g., witness derived for any pure state
\item Variance-based criteria, e.g., the well known criterion\\ ${\rm var}(x_1-x_2)+{\rm var}(p_1+p_2)>2.$
\item Multipartite case
\end{itemize}
\end{itemize}

\item Entanglement measures (How much is it entangled?)
\begin{itemize}
\item General quantum operation
\item Local operations and classical communication (LOCC)
\item Entanglement of formation
\item Concurrence
\item Entanglement of distillation
\item Bound entanglement (prove that it cannot be distilled to singlets, and its overlap with singlets is small)
\item Requirements for entanglement measures
\item Negativity
\end{itemize}

\item No-cloning theorem and related issues
\begin{itemize}
\item No cloning theorem
\item Measurement problem
\item Quantum teleportation
\item Quantum money
\item Quantum cryptography (BB84, E91)
\item Error correction (bit flip code, phase-flip code, Shore code)
\end{itemize}

\item Paper review: D. Bouwmeester, J.-W. Pan, K. Mattle, M. Eibl, H. Weinfurter, A. Zeilinger, Experimental quantum teleportation, Nature 390, 575-579 (1997).

\item Quantum metrology
\begin{itemize}
\item Simple examples of quantum metrology
\begin{itemize}
\item Classical case: Clock arm
\item Quantum case: Single spin-1/2 particle
\item Magnetometry with the fully polarized state
\item Magnetometry with the spin-squeezed state
\item Metrology with the GHZ state
\item Dicke states
\item Interferometry with squeezed photonic states
\end{itemize}
\item Entanglement theory
\begin{itemize}
\item Multipartite entanglement
\item The spin-squeezing criterion
\end{itemize}
\item Quantum metrology using the quantum Fisher information
\begin{itemize}
\item Quantum Fisher information
\item Quantum Fisher information in linear interferometers
\item Noise and imperfections
\end{itemize}
\end{itemize}

\item Paper review: Event-Ready Bell Test Using Entangled Atoms Simultaneously Closing Detection and Locality Loopholes, Wenjamin Rosenfeld, Daniel Burchardt, Robert Garthoff, Kai Redeker, Norbert Ortegel, Markus Rau, and Harald Weinfurter
Phys. Rev. Lett. 119, 010402 (2017).

\item Semidefinite programming
\begin{itemize}
\item Linear programming
\item Semidefinite programming
\item Solvable vs. not solvable by SDP
\item Basic optimization tasks solvable by SDP, for instance,
\begin{itemize}
\item $N$-representability problem (on small systems)
\item Maximum of an operator expectation value for a set given by constraints
\item Maximum of an operator expectation value for PPT states
\item Minimizing an expression given as the sum of expectation value squares
\end{itemize}
\end{itemize}
 
\end{itemize}

Literature:

\begin{itemize}

\item I. Bengtsson and K. Zyczkowski, Geometry of Quantum States, An Introduction to Quantum Entanglement, Cambridge University Press, 2009. 

\item R. Horodecki, P. Horodecki, M. Horodecki, and K. Horodecki, Quantum entanglement, Rev. Mod. Phys. 81, 865 (2009).

\item O. Gühne and G. Tóth, Entanglement detection, Phys. Rep. 474, 1 (2009); arxiv:0811.2803.

\item G. Tóth and I. Apellaniz, Quantum metrology from a quantum information science perspective, J. Phys. A: Math. Theor. 47, 424006 (2014), special issue "50 years of Bell's theorem"; arxiv:1405.4878.

\end{itemize}
\end{document}  


























